Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-10-16 2016-Weak-FH-Bisim/WeakBisim.tex
type thm 11
typo
typos again

\begin{theorem}[Congruence]\label{thm-congr-weq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two weak FH-bisimilar pNets such that 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are weak FH-bisimilar.
\end{theorem}
 \begin{proof}  (Sketch of proof) Consider $Q$ weak FH-bisimilar to $Q'$.  It means, there exist a FH-bisimulation $\mathcal{R}_{Q,Q'}$ relating the two pNets $Q$ and $Q'$. We define a relation $\mathcal{R}$ relating states of $P[Q]_{j_0}$ with states of $P[Q']_{j_0}$: 
\[\mathcal{R} = \{(\ostate{S_P\uplus S_Q},\ostate{S_P\uplus S'_Q}, \Pred_{Q,Q'})|\,(s_Q,s_{Q'}, \Pred_{Q,Q'})\in\mathcal{R}_{Q,Q'}\}\]

To prove weak FH-bisimulation of $\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$, we consider  an open transition of $\pNet[\pNetQ]_{j_0}$, and an equivalent state of $\pNet[\pNetQ']_{j_0}$, and we try to find a family of WOT of 	$\pNet[\pNetQ']_{j_0}$ that simulates it.
Consider an OT of  $\pNet[\pNetQ]_{j_0}$ it is of the form (notations introduced to prepare the decomposition):
\[
\pNet[\pNetQ]_{j_0}\models\openrule
	{
		\beta_j^{j\in( J_P\uplus J_Q)}, 
		\Pred_P\land \Pred_Q,  
		\Post_P\uplus \Post_Q}
	{\ostate{S_P\uplus S_Q} \OTWeakarrow {\alpha}
		\ostate{S'_P\uplus S'_Q}
}
\]

@@

By definition of $\mathcal{R}$ we have 
$(s_Q,s_{Q'}, \Pred_{Q,Q'})\in\mathcal{R}_{Q,Q'}$. And thus, by definition of weak FH-bisimulation, there exist a family of weak open transitions $WOT_{x}$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_Q}, \Pred_{Q' x},\Post_{Q' x}}
         {\ostate{S_{Q'}}\OTWeakarrow {\alpha_{x}} \ostate{S'_{Q' x}}}
\end{mathpar}

@@

\[\forall x.\, (S'_P\uplus S'_Q,S'_P\uplus S'_{Q' x}, \Pred_{Q,Q' x})\in\mathcal{R}\]
which is trivially verified, and
\begin{equation*}
\begin{split}
\Pred_{Q,Q'}\land \Pred_P\land \Pred_Q \implies &
\Big(\bigvee_{x\in X}
(\forall j\in J_Q.\, \vis{\beta_j}=\gamma_{j x}\land \forall j\in J_P.\, \vis{\beta_j}=\vis{\beta_j}))\Rightarrow\\
& (\Pred_P\land \Pred_{Q' x}\land\\ &~~\alpha=\alpha \land \Pred_{Q,Q' x}\subst{\Post_P\uplus\Post_{Q' x}\uplus \Post_{Q}}) \Big)
\end{split}
\end{equation*}
We conclude by observing that $\Post_P$ has no effect on variables of $Q$ and $Q'$, and thus on $\Pred_{Q,Q' x}$
\qed

>>>>

\begin{theorem}[Congruence]\label{thm-congr-weq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two weak FH-bisimilar pNets such that 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are weak FH-bisimilar.
\end{theorem}
 \begin{proof}  Consider $Q$ weak FH-bisimilar to $Q'$.  It means that there exist a FH-bisimulation $\mathcal{R}_{Q,Q'}$ relating the two pNets $Q$ and $Q'$. We define a relation $\mathcal{R}$ relating states of $P[Q]_{j_0}$ with states of $P[Q']_{j_0}$: 
\[\mathcal{R} = \{(\ostate{S_P\uplus S_Q},\ostate{S_P\uplus S_{Q'}}, \Pred_{Q,Q'})|\,(S_Q,S_{Q'}, \Pred_{Q,Q'})\in\mathcal{R}_{Q,Q'}\}\]

To prove weak FH-bisimulation of $\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$, we consider  an open transition $OT$ of $\pNet[\pNetQ]_{j_0}$, and an equivalent state of $\pNet[\pNetQ']_{j_0}$, and we try to find a family of WOT of 	$\pNet[\pNetQ']_{j_0}$ that simulates $OT$.
Consider an OT of  $\pNet[\pNetQ]_{j_0}$ it is of the form (notations introduced to prepare the decomposition):
\[
\pNet[\pNetQ]_{j_0}\models\openrule
	{
		\beta_j^{j\in( J_P\uplus J_Q)}, 
		\Pred_P\land \Pred_Q,  
		\Post_P\uplus \Post_Q}
	{\ostate{S_P\uplus S_Q} \OTWeakarrow {\alpha}
		\ostate{S'_P\uplus S'_Q}
}
\]

@@

By definition of $\mathcal{R}$ we have 
$(S_Q,S_{Q'}| \Pred_{Q,Q'})\in\mathcal{R}_{Q,Q'}$. And thus, by definition of weak FH-bisimulation, there exist a family of weak open transitions $WOT_{x}$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_Q}, \Pred_{Q' x},\Post_{Q' x}}
         {\ostate{S_{Q'}}\OTWeakarrow {\alpha_{x}} \ostate{S'_{Q' x}}}
\end{mathpar}

where  
\[\forall x.\, (S'_Q,S'_{Q' x}| \Pred_{Q,Q' x})\in\mathcal{R}_{Q,Q'}\]
and
\begin{equation*}
\begin{split}
\Pred_{Q,Q'}\land \Pred_Q \implies &
\Big(\bigvee_{x\in X}
(\forall j\in J_Q.\, \vis{\beta_j}=\gamma_{j x})\Rightarrow (\Pred_{Q_x}\land\\ &~~\alpha_Q=\alpha_x \land \Pred_{Q,Q' x}\subst{\Post_{Q' x}\uplus \Post_{Q}}) \Big)
\end{split}
\end{equation*}

Composing the OT of $P$ with the WOTs of $Q'$ by Lemma~\ref{lem-Weakcompose1} we obtain:
\[
\pNet[\pNetQ']_{j_0}\models\openrule
	{
		\vis{\beta_j^{j\in( J_P}}\uplus\gamma_{j x}^{j\in J_Q}, 
		\Pred_P\land \Pred_{Q' x},  
		\Post_P\uplus \Post_{Q' x}}
	{\ostate{S_P\uplus S_{Q'}} \OTWeakarrow {\alpha}
		\ostate{S'_P\uplus S'_{Q' x}}
}
\]
with $\bigvee_{x\in X}
(\forall j\in J_Q.\, \vis{\beta_j}=\gamma_{j x} \implies \alpha_Q=\alpha_x$ that ensures that the open transitions can be recomposed when the OT fires.

\[\forall x.\, (S'_P\uplus S'_Q,S'_P\uplus S'_{Q' x}| \Pred_{Q,Q' x})\in\mathcal{R}\]
which is trivially verified, and
\begin{equation*}
\begin{split}
\Pred_{Q,Q'}\land \Pred_P\land \Pred_Q \implies &
\Big(\bigvee_{x\in X}
(\forall j\in J_Q.\, \vis{\beta_j}=\gamma_{j x}\land \forall j\in J_P.\, \vis{\beta_j}=\vis{\beta_j}))\Rightarrow\\
& (\Pred_P\land \Pred_{Q' x}\land\\ &~~\alpha=\alpha \land \Pred_{Q,Q' x}\subst{\Post_P\uplus\Post_{Q' x}\uplus \Post_{Q}}) \Big)
\end{split}
\end{equation*}
We conclude by observing that $\Post_P$ has no effect on variables of $Q$ and $Q'$, and thus on $\Pred_{Q,Q' x}$
\qed

 \begin{proof}   Consider $P$ weak FH-bisimilar to $P'$.  It means, there exist a FH-bisimulation $\mathcal{R}_{P,P'}$ relating the two pNets $P$ and $P'$. We define a relation $\mathcal{R}$ relating states of $P[Q]_{j_0}$ with states of $P'[Q]_{j_0}$: 
\[\mathcal{R} = \{(\ostate{S_P\uplus S_Q},\ostate{S_{P'}\uplus S_Q}, \Pred_{P,P'})|\,(S_P,S_{P'}, \Pred_{P,P'})\in\mathcal{R}_{P,P'}\}\]

To prove weak FH-bisimulation of $\pNet[\pNetQ]_{j_0}$ and 
	$\pNet'[\pNetQ]_{j_0}$, we consider  an open transition $OT$ of $\pNet[\pNetQ]_{j_0}$, and an equivalent state of $\pNet'[\pNetQ]_{j_0}$, and we try to find a family of WOT of 	$\pNet'[\pNetQ]_{j_0}$ that simulates $OT$.
Consider an OT of  $\pNet[\pNetQ]_{j_0}$ it is of the form (notations introduced to prepare the decomposition):
\[
\pNet[\pNetQ]_{j_0}\models\openrule
	{
		\beta_j^{j\in( J_P\uplus J_Q)}, 
		\Pred_P\land \Pred_Q,  
		\Post_P\uplus \Post_Q}
	{\ostate{S_P\uplus S_Q} \OTWeakarrow {\alpha}
		\ostate{S'_P\uplus S'_Q}
}
\]

By definition of $\mathcal{R}$ we have 
$(S_P,S_{P'}, \Pred_{Q,Q'})\in\mathcal{R}_{Q,Q'}$. And thus, by definition of weak FH-bisimulation, there exist a family of weak open transitions $WOT_{x}$:
 \begin{mathpar}
    \openrule
         {
           \gamma_{j x}^{j\in J_Q}, \Pred_{Q' x},\Post_{Q' x}}
         {\ostate{S_{Q'}}\OTWeakarrow {\alpha_{x}} \ostate{S'_{Q' x}}}
\end{mathpar}

@@

\[\forall x.\, (S'_P\uplus S'_Q,S'_P\uplus S'_{Q' x}, \Pred_{Q,Q' x})\in\mathcal{R}\]
which is trivially verified, and
\begin{equation*}
\begin{split}
\Pred_{Q,Q'}\land \Pred_P\land \Pred_Q \implies &
\Big(\bigvee_{x\in X}
(\forall j\in J_Q.\, \vis{\beta_j}=\gamma_{j x}\land \forall j\in J_P.\, \vis{\beta_j}=\vis{\beta_j}))\Rightarrow\\
& (\Pred_P\land \Pred_{Q' x}\land\\ &~~\alpha=\alpha \land \Pred_{Q,Q' x}\subst{\Post_P\uplus\Post_{Q' x}\uplus \Post_{Q}}) \Big)
\end{split}
\end{equation*}
We conclude by observing that $\Post_P$ has no effect on variables of $Q$ and $Q'$, and thus on $\Pred_{Q,Q' x}$
\qed
\end{proof}
